---
title: "Guide to Macrosystems EDDIE Shiny applications"
output: 
  html_document:
    code_folding: "hide"
    toc: true
runtime: shiny
---

```{r setup, include=FALSE}
library(ggplot2)
library(plotly)
library(DT)
library(hover)
library(shinyBS)
knitr::opts_chunk$set(echo = TRUE)
module_text <- read.csv("app/data/module_text.csv", row.names = 1, header = FALSE)
chla <- readr::read_csv("app/data/neon/BARC_chla_microgramsPerLiter.csv", show_col_types = FALSE)
neon_sites_df <- read.csv("app/data/neon_sites.csv")
neon_sites_df$long <- round(neon_sites_df$long, 3)
neon_sites_df$lat <- round(neon_sites_df$lat, 3)
neon_sites_df <- neon_sites_df[which(neon_sites_df$siteID %in% c("CRAM", "BARC", "PRPO", "LIRO", "PRLA")), ]
# Tab names for updating buttons
tab_names <- read.csv("app/data/tab_names.csv", fileEncoding = "UTF-8-BOM")
rownames(tab_names) <- tab_names[, 1]

```


# Contents
1. File structure
2. Updating text
3. Updating plots
4. Updating tables
5. Publishing on shinyapps.io
6. Page buttons
7. Saving progress
8. Uploading progress
9. Rendering report

## 1. File structure
Each of the shiny apps follows the same directory structure:


```
module7
    ¦--app
    ¦   ¦--R                        # R functions to be sourced
    ¦   ¦   ¦--NP_model.R
    ¦   ¦   ¦--create_np_input.R
    ¦   ¦   ¦--etc...
    ¦   ¦--data                     # Shiny specific directory for data files (e.g. *.csv, *.txt, *.rds, etc.)
    ¦   ¦   ¦--data_collection_method.csv 
    ¦   ¦   ¦--...
    ¦   ¦--icons                    # icon used for Leaflet maps
    ¦   ¦   ¦--water-icon.png
    ¦   ¦   ¦--...
    ¦   ¦--www                      # Shiny specific directory for images (e.g. *.png, *.jpg, etc.)
    ¦   ¦   ¦--EnKF_figure.png
    ¦   ¦   ¦--...
    ¦   ¦--report.Rmd               # R markdown which gets rendered into a Word doc with the students answeres
    ¦   ¦--report-style.docx        # Word doc which contains styles used for report.Rmd
    ¦   ¦--global.R                 # Loads packages & non-reactive data on start-up (e.g. module text, questions, etc.)
    ¦   ¦--server.R                 # Takes inputs from UI and returns it as outputs to UI
    ¦   ¦--ui.R                     # User interface - generates the html code for the text, images, inputs, etc.
    ¦--README.md                    # Readme file for GitHub
    ¦--apt.txt                      # Required for Binder
    ¦--runtime.txt                  # Required for Binder
    ¦--install.R                    # Required for Binder
```

The `app` directory contains all the files necessary for the Shiny app.

#### `data`
This directory contains all the data files. CSV files with module text, tab names, module text, NEON site data and NOAA data.

#### `icons`
This directory contains the icons that are used for the leaflet map of the NEON sites.

#### `www`
This directory contains all the supporting image files seen within the app.

#### `global.R`
This R script loads the libraries and reads in most of the data files which are then available across all Shiny applications. For example, on shinyapps.io there can be multiple workers on one instance, or in layperson's terms, there can be multiple shiny apps running from one computer, so rather than have each shiny app run this script this is run when the computer starts up and all the packages are loaded, common variables (such as the module text, NEON site data, etc.) and R functions are loaded into the environment and shared across each of the apps.

If you add a new package dependency, you would add this here.

#### `report.Rmd`
This is the Rmarkdown file that is used to generate the final report. This will be explained in more detail in the Rendering report section.

#### `report-style.docx`
This is a Word document that contains the styles used when rendering the report.

#### `server.R`
This is the R file where code

#### `ui.R`
This R file contains the R functions to generate the HTML code which is rendered in the internet browser (with javascript). It contains the id's (`inputId`) of each interactive element (e.g. sliders, buttons, plots, datatables etc.) which is linked in `server.R`.

## 2. Updating text
There are two places you will need to look to edit the text:
```
module7                                                       
    ¦--app
        ¦--data 
        ¦   ¦--module_text.csv
        ¦--ui.R
```

1. Open the `ui.R` script, launch the app in your browser for reference then locate the text within the script.
2. Text is either written directly within the code e.g. `p("Some text")` or,
3. It is sourced from pre-loaded data frame e.g. `p(module_test["some_text",])`

If it is pre-loaded in a dataframe, there will be a corresponding csv file in the `data` directory and it is loaded within the `global.R` script. The reasoning behind this was to reduce the amount of text within the shiny app, and also to help facilitate easy update of text and transfer common text between modules. For example, a lot of forecasting definitions are in the csv file which is replicated across the modules.

Below is the example Shiny text with the corresponding Shiny output (blue). There is a dataframe with daily chlorophyll-a data for Lake Barco.
```{r}
# R Shiny code
fluidRow(
  hr(), # Adds horizontal line
  column(6, style = "background-color:#DCF6F9;",
         h5("{Rendered Shiny output}"),
         h3("Presentation"),
         p("The presentation accompanying this module reviews the forecast cycle, introduces the process of using ecological data to improve forecasts (data assimilation), and explains how the amount of uncertainty in observations and model predictions as well as the frequency of observations affect data assimilation."),
         p("What is data assimilation?"),
         tags$ul(
           tags$li(module_text["data_assimilation", ])
           )
         ),
  hr() # Adds horizontal line
  )
```

## 3. Updating plots
Within the shiny apps, plots are either formatted as plotly plots (interactive plots) or standard rendered plots (no interactivity). The majority of plots are rendered as plotly plots.

```{r}
# Simple plotting code
p <- ggplot() +
  geom_point(data = chla, aes(Date, V1)) +
  ggtitle("")

# server.R - Shiny code
# Static plot
renderPlot({
  p +
    ggtitle("Static plot - Chl-a at Barco") +
    theme_bw()
})

renderPlotly({
  p1 <- p +
    ggtitle("Interactive plot") +
    theme_bw()
  ggplotly(p1)
})

```

All plots are built with ggplot so for adjusting any features (e.g. axis labels, etc.), regular ggplot functions apply (`xlab()`, `theme_classic()`.



## 4. Updating tables
All tables within the Macrosystems EDDIE Shiny apps are rendered using the datatable package, but for R Shiny the package is named `DT`.

```{r}
# R Shiny code
renderDT(chla)
```
Data tables are really powerful pieces of javascript and the basic rendering contains many key cool features e.g. search bar, ordering of columns, pages of data etc. But this can be overkill in most instances and can give users too much flexibility so therefore it is needed to adjust the rendering settings.

Data tables can quickly get complicated and the documentation is thorough, but keyed towards Javascript programmers. There are X main types of data tables we use:
a. With pages of data
b. Selectable rows
c. Static tables

### a. Data table - pages of data
Here is a data table with pages of data, but with row names switched off and column names added, and with the search bar disabled and limited to 4 rows per page.

This is used in Objective 2 - Explore data.
```{r}
# R Shiny code
renderDT(chla,
  rownames = FALSE, colnames = c("Date", "Chl-a"),
  options = list(searching = FALSE, pageLength = 4)
  )
```

### b. Data table - Selectable rows
By default, within data table, rows can be selected. Below we will render the data table

```{r}
actionButton(inputId = "show_tab", "Show table")
renderDT({
  validate(
    need(input$show_tab > 0, message = "Click 'Show table'")
  )
  neon_sites_df[, c(1:2)]
  },
    selection = "single", rownames = FALSE,
    options = list(dom = 't') # Renders just the table
  )
```
\

## 6. Page buttons
These are actually a bit trickier than you might think, but this could be because I used a very "hacky" way to do this as I was surprised this feature 

```{r warning = FALSE}
nav_bg <- "#DDE4E1"
nav_butt <- "#31ED92"
nav_txt <- "#000000" # white = #fff; black = #000000

navbarPage(title = "Module 7: Using Data to Improve Ecological Forecasts",
               position = "static-top", id = "maintab",
           tabPanel(title = tab_names["mtab1", 2], value = "mtab1",
                    img(src = "project-eddie-banner-2020_green.png", height = 100,
                     width = 1544, top = 5),
                    h1("Overview")
                    ),
           tabPanel(title = tab_names["mtab2", 2], value = "mtab2",
                        img(src = "project-eddie-banner-2020_green.png", height = 100,
                            width = 1544, top = 5),
                    h1("Introduction")
           ),
           tabPanel(title = tab_names["mtab3", 2], value = "mtab2",
                        img(src = "project-eddie-banner-2020_green.png", height = 100,
                            width = 1544, top = 5),
                    h1("Presentation")
           ),
           fluidRow(
            column(5, align = "center",
                   br(),
                   hover_action_button(
                     inputId = "prevBtn1",
                     label = "< Previous",
                     button_animation = "glow",
                     style = paste0("color: ", nav_txt, "; background-color: ", nav_butt, "; border-color: #00664B; padding:15px; font-size:22px;")
                   ),
                   bsTooltip("prevBtn1", title = "Navigate to previous tab", placement = "left", trigger = "hover"),
                   br(), br()
                   
            ),
            column(2, align = "center",
                   br(),
                   tags$style(type="text/css", paste0("#download_answers {background-color:#579277;color: white; padding:15px; font-size:18px;}")),
                   hover_download_button(outputId = "download_answers",
                                         label = "Download user input",
                                         class = "butt1",
                                         button_animation = "glow"),
                   br(), br()
            ),
            column(5, align = "center",
                   br(),
                   use_hover(popback = TRUE),
                   hover_action_button(
                     inputId = "nextBtn1",
                     label = "Next >",
                     button_animation = "glow",
                     style = paste0("color: ", nav_txt, "; background-color: ", nav_butt, "; border-color: #00664B; padding:15px; font-size:22px;")
                   ),
                   bsTooltip("nextBtn1", title = "Navigate to next tab", placement = "right", trigger = "hover"),
                   br(), br()                   # )

                   )
            )
          )


```


## 5. Publishing on shinyapps.io
## 7. Saving progress
## 8. Uploading progress
## 9. Rendering report

